---
title: "MA678 Homework 2"
author: "Runci Hu"
date: "9/20/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
remotes::install_github("avehtari/ROS-Examples", subdir = "rpackage")
library(rosdata)
data(kidiq)
library(rstanarm)
```

## 11.5 
*Residuals and predictions*: The folder `Pyth` contains outcome $y$ and predictors $x_1$, $x_2$ for 40 data points, with a further 20 points with the predictors but no observed outcome. Save the file to your working directory, then read it into R using `read.table()`.

### (a) 
Use R to fit a linear regression model predicting $y$ from $x_1$, $x_2$, using the first 40 data points in the file. Summarize the inferences and check the fit of your model.

```{r}
data <- "https://raw.githubusercontent.com/avehtari/ROS-Examples/master/Pyth/pyth.txt" 
pyth <- read.table(file = data, header = TRUE)
reg1 <- lm(y~x1+x2, data = pyth)
summary(reg1)
```

### (b) 
Display the estimated model graphically as in Figure 10.2

```{r}
library(ggplot2)
ggplot1 <- ggplot(reg1, aes(x=x1, y=y))+geom_point()+geom_smooth(method="lm") + labs(title="y~x1", x="x1", y="y")
ggplot2 <- ggplot(reg1, aes(x=x2, y=y))+geom_point()+geom_smooth(method="lm") + labs(title="y~x2", x="x2", y="y")
ggplot1
ggplot2
```

### (c) 
Make a residual plot for this model. Do the assumptions appear to be met?

```{r}
plot(reg1,1)

#No, they don't meet
```




### (d) 
Make predictions for the remaining 20 data points in the file. How confident do you feel about these predictions?

```{r}
data_d<-data.frame(x1=pyth$x1[41:60],x2=pyth$x2[41:60])
predict(reg1,data_d)
##yes, i am confident about this prediction. 
```


## 12.5 
*Logarithmic transformation and regression*: Consider the following regression:
$$\log(\text{weight})=-3.8+2.1 \log(\text{height})+\text{error,} $$
with errors that have standard deviation 0.25. Weights are in pounds and heights are in inches.

### (a) 
Fill in the blanks: Approximately 68% of the people will have weights within a factor of __-1.28____ and __1.28____ of their predicted values from the regression.

### (b) 
Using pen and paper, sketch the regression line and scatterplot of log(weight) versus log(height) that make sense and are consistent with the fitted model. Be sure to label the axes of your graph.
```{r}
x <- log(earnings$height)
y <- log(earnings$weight)
reg2 <- lm(y~x)
plot(x,y, xlab="log height", ylab="log weight")
abline(coef(reg2)[1], coef(reg2)[2], col="red")
```


## 12.6 
*Logarithmic transformations*: The folder `Pollution` contains mortality rates and various environmental factors from 60 US metropolitan areas. For this exercise we shall model mortality rate given nitric oxides, sulfur dioxide, and hydrocarbons as inputs. this model is an extreme oversimplication, as it combines all sources of mortality and does not adjust for crucial factors such as age and smoking. We use it to illustrate log transformation in regression.  

### (a) 
Create a scatterplot of mortality rate versus level of nitric oxides. Do you think linear regression will fit these data well? Fit the regression and evaluate a residual plot from the regression.

```{r}
library(rosdata)
data("pollution")
y <- pollution$mort
x <- pollution$nox

plot(x,y)
reg3 <- lm(y~x, data=pollution)
print(reg3)
abline(coef(reg3)[1], coef(reg3)[2],col="red")

plot(reg3,1)
##Not fit well
```

### (b) 
Find an appropriate reansformation that will result in data more appropriate for linear regression. Fit a regression to the transformed data and evaluate the new residual plot.

```{r}
reg4 <- lm(log(y)~log(x), data=pollution)

plot(log(x),log(y))
abline(lm(log(y)~log(x), data=pollution),col="red")

plot(reg4, 1)
```

### (c) 
Interpret the slope coefficient from the model you chose in (b)
they are symmetry
For each 1% difference in mortality rate, 
the prediction difference in log transformation is 0.015%

### (d) 
Now fit a model predicting mortality rate using levels of nitric oxides, sulfur dioxide, and hydrocarbons as inputs. Use appropriate transformation when helpful. Plot the fitted regression model and interpret the coefficients.

```{r}
a <- pollution$mort
b <- pollution$nox
c <- pollution$hc
d <- pollution$so2
fit <- lm(formula = log(a) ~ log(b)+log(c)+log(d), data=pollution, refresh=0)
fit
bmean<-mean(log(pollution$nox))
cmean<-mean(log(pollution$hc))
dmean<-mean(log(pollution$so2))
plot(log(a)~log(b))
abline(coef(fit)[1]+cmean*coef(fit)[3]+dmean*coef(fit)[4], coef(fit)[2], col="red")
plot(log(a)~log(c))
abline(coef(fit)[1]+bmean*coef(fit)[2]+dmean*coef(fit)[4], coef(fit)[3], col="yellow")
plot(log(a)~log(d))
abline(coef(fit)[1]+bmean*coef(fit)[2]+cmean*coef(fit)[3], coef(fit)[4], col="blue")
##Using log-log transformation, the slope coefficient means that 
##for each 1% difference in nox, the predicted difference in mort is 0.06%
##for each 1% difference in hc, the predicted difference in mort is negative 0.06%
##for each 1% difference in so2, the predicted difference in mort is 0.01%


```

### (e) 
Cross validate: fit the model you chose above to the first half of the data and then predict for the second half. You used all the data to construct the model in (d), so this is not really cross validation, but it gives a sense of how the steps of cross validation can be implemented.

```{r}
fit2 <- lm(log(a)[1:30] ~ log(b)[1:30]+log(c)[1:30]+log(d)[1:30], data=pollution)
secondhalf <- data.frame(b1=log(b)[31:60], c1=log(c)[31:60], d1=log(d)[31:60])
predict <- predict(fit2, secondhalf)
plot(predict, predict-log(a)[31:60])
```

## 12.7 
*Cross validation comparison of models with different transformations of outcomes*: when we compare models with transformed continuous outcomes, we must take into account how the nonlinear transformation warps the continuous outcomes. Follow the procedure used to compare models for the mesquite bushes example on page 202.

### (a) 
Compare models for earnings and for log(earnings) given height and sex as shown in page 84 and 192. Use `earnk` and `log(earnk)` as outcomes.

```{r}
e <- earnings$earnk
h <-earnings$height
m <- earnings$male
reg5 <- stan_glm(e~h+m, data=earnings, refresh=0, subset = earn>0)
reg6_log <- stan_glm(log(e)~h+m, data=earnings, refresh=0, subset = earn>0)

(loo_compare(loo(reg5), loo(reg6_log))) ##comparing 2 models with loo, 
                                        ##see textbook
```

### (b) 
Compare models from other exercises in this chapter.
```{r}
## compare models in 12.14
fit_14 <- stan_glm(formula = weight ~ group + diam1*diam2 + diam1/diam2 + canopy_height + total_height + density, data=mesquite, refresh=0)
fit_14_log <- stan_glm(formula = log(weight) ~ log(diam1*diam2) + log(diam1/diam2) +
       log(canopy_height) + log(total_height) + log(density) + group, data=mesquite, refresh=0)
(loo_compare(loo(fit_14), loo(fit_14_log)))
```


## 12.8 
*Log-log transformations*: Suppose that, for a certain population of animals, we can predict log weight from log height as follows:  

* An animal that is 50 centimeters tall is predicted to weigh 10 kg.

* Every increase of 1% in height corresponds to a predicted increase of 2% in weight.

* The weights of approximately 95% of the animals fall within a factor of 1.1 of predicted values.

### (a) 
Give the equation of the regression line and the residual standard deviation of the regression.

ln(10)=Intercept+2*ln(50) Int=-5.52146
ln(1.1)/2=0.048
log(weight)=-5.52146+2*log(height), residual=0.048

### (b) 
Suppose the standard deviation of log weights is 20% in this population. What, then, is the $R^{2}$ of the regression model described here? 

log(1.1)/2=0.048
1-(0.048^2/0.2^2)
0.94

## 12.9 
*Linear and logarithmic transformations*: For a study of congressional elections, you would like a measure of the relative amount of money raised by each of the two major-party candidates in each district. Suppose that you know the amount of money raised by each candidate; label these dollar values $D_i$ and $R_i$. You would like to combine these into a single variable that can be included as an input variable into a model predicting vote share for the Democrats. Discuss the advantages and disadvantages of the following measures:  

### (a) 
The simple difference, $D_i - R_i$

Advantage: see the difference between values, 
           using the intercept to find difference when =0

### (b) 
The ratio, $D_i / R_i$

Easily see the ratio difference between parties of D and R
the shortage is that this outcome is not straightforward for the difference. 

### (c) 
The difference on the logarithmic scale, $\log D_i - \log R_i$

It indicates the relationship between log vote share and log value difference.
The log transformation creates a better fitted value 
for the model with large inputs. 
The disadvantage is that the difference will become smaller

### (d) 
The relative proportion, $D_{i}/(D_{i}+R_{i})$. 

It shows the ratio of vote share for D in the total vote share. 
With this ratio, we can also calculate the ratio for R. 


## 12.11
*Elasticity*: An economist runs a regression examining the relations between the average price of cigarettes, $P$, and the quantity purchased, $Q$, across a large sample of counties in the United  States, assuming the functional form, $\log Q=\alpha+\beta \log P$. Suppose the estimate for $\beta$ is 0.3.  Interpret this coefficient. 

exp(0.3)=1.34986
Every 1% increase in quantity corresponds to 
an increase of 34.99% in prediciated average price of cigarettes.


## 12.13
*Building regression models*: Return to the teaching evaluations data from Exercise 10.6. Fit regression models predicting evaluations given many of the inputs in the dataset. Consider interactions, combinations of predictors, and transformations, as appropriate. Consider several  models, discuss in detail the final model that you choose, and also explain why you chose it rather than the others you had considered. 

```{r}
reg7 <- stan_glm(eval~beauty+female+age+female*age,data=beauty,refresh=0)
reg8 <- stan_glm(log(eval)~beauty+female+age+female*age,data=beauty,refresh=0)

(loo_compare(loo(reg7),loo(reg8)))

##I would choose reg 8 that has log transformation of eval. 
##Because 0 is better than a negative number. 
```


## 12.14
Prediction from a fitted regression: Consider one of the fitted models for mesquite leaves, for example `fit_4`, in Section 12.6. Suppose you wish to use this model to make inferences about the average mesquite yield in a new set of trees whose predictors are in data frame called  new_trees. Give R code to obtain an estimate and standard error for this population average. You do not need to make the prediction; just give the code. 

```{r}
fit_14 <- stan_glm(formula = log(weight) ~ group + log(diam1*diam2) + log(diam1/diam2) + log(canopy_height) + log(total_height) + log(density), data=mesquite, refresh=0)
fit_14
```
